Minimizing Time Required For Graduation

### 2010-12-11 paper/paper.tex
Rewrite/edit sections 1,2....again.

\abstract{The Bachelor's Degree is designed to take four years to complete,
however, many students fail to complete it this amount of time. One 
factor contributing to this is poor planning when registering for classes.
We present an algorithm to assist students in
constructing a schedule that works towards completing a degree in the shortest
amount of time possible.}
\pagebreak

\section{Introduction} \subsection{Background} Registration can be a stressful
time for every student at the University of Washington. Without good
planning, students can make poor registration choices resulting in more quarter
being needed to graduate than otherwise necessary. Picking the correct class
involves considering quarterly class offerings, class credits, the prerequisites
of a class, time conflicts, and how to fulfill future prerequisites. The
University of Washington registration system already provides a tool to aid
students, schedule finder. Schedule finder however, only takes into account time
conflicts in one quarter, and does not consider what classes are best for
minimizing graduation time.

A student's graduation time may be extended because of his or her inability to
take a class that has limited offerings during that particular year.

\subsection{Goal} Our goal is to provide an automated system that gives quarter
by quarter schedules until a student graduates. As input our system will take in
the current set of classes a student has taken and the preferred credit hours
per quarter. Additional input will be the current and past university time
schedules. As output our system will provide a class schedule that minimizes the
number of quarters until graduation.

@@

\section{Problem Description} To make our examples concrete we will focus on the
classes offered within the Math department, although our algorithm and model can
be extended to include over departments.

Every quarter the Math department offers a subset of courses from all of the
available courses in the department. Each course is offered during a certain set
time periods. There are inevitably time conflicts between courses that a student
could take because there are far more courses offered than time periods during
the regular five-day week  To graduate with a B.S. in Mathematics there are
certain classes that are required and then there are elective requirements too.
Additionally many classes have prerequisites which much be satisfied before they
can be taken. In the case of math courses, there can be multiple, exclusive
prerequisite sets that independently satisfy the requirements for a class.

@@

\subsection{Definitions} We will now formalize some terms used in our paper.
\begin{itemize} \item A {\it course} refers to any course offered by the
University of Washington. A course is identified by its department name and
number, for example {\it MATH309}. Each course is also worth a specific amount
of credits. \item We refer to a specific instance of a course as a {\it course
offering}. Each course offering additionally has associated with it the time of
the day at which the course meets, the quarter it is being offered, and the
section number of the offering. Throughout this paper, we will refer to quarters
by their time schedule abbreviates: {\it F} for Fall Quarter, {\it W} for Winter
Quarter, {\it Sp} for Spring Quarter, and {\it S} for Summer Quarter.
\end{itemize}

@@

\subsection{Representing Prerequisites} Initially, we hoped to represent course
prerequisites using a directed acyclic graph. Each node was a course, and edges
existed between nodes $i$ and $j$ if and only if $i$ was a prerequisite for $j$.
The DAG relationship works well when courses have single requirements. It
becomes difficult, however, to track whether requires are fulfilled with courses
that have multiple prerequisites, or worse courses that have alternative
prerequisites. We decided that it was would be feasible to model prerequisites
using a graph.
\begin{figure} [ht] \begin{center}
\includegraphics[scale=0.4]{prereq_logic_example} \end{center}
\caption{Evaluation of a rule} \label{logic_ex} \end{figure}

@@

\section{Picking Classes Per Quarter} Since determining an entire schedule at
once is intractable we adapted the rolling horizons approach which uses
a heuristic to pick the best possible schedule for a quarter. With a good
heuristic we will be able to chose classes that lead to a minimized graduation
time. Since we cannot analytically determine graduation time based on classes
chosen in a single quarter, our heuristic tries to give high value to classes
that would delay graduation is taken later rather than sooner.

>>>>

\abstract{The Bachelor's Degree is designed to take four years to complete;
however, many students fail to complete it in this time frame. One 
factor contributing to this is poor planning when registering for classes.
We present an algorithm to assist students in constructing a schedule 
that works towards completing a degree in the shortest amount of time 
possible. The algorithm presented uses heurisitics to attempt to pick the best
class set each quarter, working towards a complete schedule quartery by 
quarter.}
\pagebreak

\section{Introduction} \subsection{Background} Registration can be a stressful 
time for students at the University of Washington. Without proper planning, 
a student can run the risk of making poor registration choices, which can 
result in the student taking more quarters than otherwise necessary. However, 
choosing the right schedule is complex and there are multiple variables to 
consider. Picking classes involves the consideration of quarterly class 
offerings, class credits, prerequisites, time conflicts, and whether or not 
classes will fulfill future prerequisites. All of these factors can affect a 
student’s graduation time, especially if that class has limited offerings 
during a particular year. The University of Washington registration system 
already provides a tool to aid students: schedule finder. However, schedule 
finder only takes into account time conflicts per quarter, and not what 
classes are best for minimizing graduation time.

\subsection{Goal} Our goal is to provide an automated system that gives quarter
by quarter schedules up until graduates. As input our system will take in
the current set of classes a student has taken and the preferred credit hours
per quarter. Additional input will be the current and past university time
schedules. As output our system will provide a class schedule that minimizes the
number of quarters until graduation.

@@

\section{Problem Description} To make our examples concrete we will focus on the
classes offered within the Math department; however, our algorithm and model can
be extended to include other departments.

Every quarter the Mathematics department offers a selection of courses; not every
course is offered every quarter. Each course is offered during a certain set
time periods. There are inevitable time conflicts among courses that a student
could take because there are far more courses offered than time periods during
the regular five-day week. Graduating with a B.S. in Mathematics requires
specific classes in addition to elective requirements.
Additionally many classes have prerequisites which much be satisfied before they
can be taken. In the case of math courses, there can be multiple, exclusive
prerequisite sets that independently satisfy the requirements for a class.

@@

\subsection{Definitions} We will now formalize some terms used in our paper.
\begin{itemize} \item A {\it course} refers to any course offered by the
University of Washington. A course is identified by its department name and
number, for example {\it MATH309}. Each course is also worth a specific amount
of credits. \item We refer to a specific instance of a course as a {\it course
offering}. Additionally, each course offering  has associated with it the time of
the day at which the course meets, the quarter it is being offered, and the
section number of the offering. Throughout this paper, we will refer to quarters
by their time schedule abbreviates: {\it F} for Fall Quarter, {\it W} for Winter
Quarter, {\it Sp} for Spring Quarter, and {\it S} for Summer Quarter.
\end{itemize}

@@

\subsection{Representing Prerequisites} Initially, we hoped to represent course
prerequisites using a directed acyclic graph. Each node was a course, and edges
existed between nodes $i$ and $j$ if and only if $i$ was a prerequisite for $j$.
The DAG relationship works well when courses have single requirements. It
becomes difficult, however, to track whether requirements are fulfilled with courses
that have multiple prerequisites, or worse courses that have alternative
prerequisites. We decided that it would be feasible to model prerequisites
using a graph.
\begin{figure} [ht] \begin{center}
\includegraphics[scale=0.4]{prereq_logic_example} \end{center}
\caption{Evaluation of a rule} \label{logic_ex} \end{figure}

@@

\section{Picking Classes Per Quarter} Since determining an entire schedule at
once is intractable we adapted the rolling horizons approach which uses
a heuristic to pick the best possible schedule for a quarter. With a good
heuristic we will be able to choose classes that lead to a minimized graduation
time. Since we cannot analytically determine graduation time based on classes
chosen in a single quarter, our heuristic tries to give high value to classes
that would delay graduation is taken later rather than sooner.
